---
title: "STAT 445 Assignment 2"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 11.27

11.27 a)
Read in the dataset.
```{r}
t11_5 <- read.table("T11-5.DAT.txt", header = F, col.names=c("sepal_length", "sepal_width", "petal_length", "petal_width", "groups"))
t11_5$groups <- factor(t11_5$groups, labels = c("setosa", "versicolor", "virginica"))
summary(t11_5)
```

```{r}
plot(as.matrix(t11_5[,c(2,4)]), xlab="", ylab="", col = as.integer(t11_5$groups), pch= as.integer(t11_5$groups))
```
Looking at the plot, it is not fair to assume a multivariate normal distribution as the cluster of data points for each group is not shaped elliptically.
c)
```{r}
library(MASS)
qda_model <- qda(groups ~ sepal_width + petal_width, data=t11_5, prior=c(1/3, 1/3, 1/3) )
qda_pred <- predict(qda_model, newdata = data.frame(sepal_width=3.5, petal_width=1.75))
xbar <- by(t11_5[,c(2,4)], t11_5$groups, colMeans )
n <- rep(50,3)
S <- by(t11_5[,c(2,4)], t11_5$groups, var)
Spool <- (S[[1]]*(n[1]-1) + S[[2]]*(n[2]-1) + S[[3]]*(n[3]-1))/(sum(n-1))
prior <- c(1/3, 1/3, 1/3)
for (i in 1:3) {
  print(log(det(S[[i]])))
}
```
The above list all the \[ ln(|S_i|)\].

d)
```{r}
ceofMat <- matrix(0, 3, 3)

for(i in 1:nrow(ceofMat)){
ceofMat[i,] <- c(-0.5*t(xbar[[i]])%*%solve(Spool)%*%xbar[[i]],
                 t(xbar[[i]])%*%solve(Spool)) 
}
ceofMat
```
The above is the matrix of coefficients for the linear discrimate formula.

e) Functins 'yval', 'dhat', and 'drawline' are taken from ldaClassificationBoundary.rmd.
```{r}
d_matrix <- matrix(0, 3, 3)
x = c(3.5, 1.75)
for (i in 1:3) {
  for (j in 1:3)
    d_matrix[i,j] = t(xbar[[i]]-xbar[[j]])%*%solve(Spool)%*%x - 0.5*t(xbar[[i]]-xbar[[j]])%*%solve(Spool)%*%(xbar[[i]]+xbar[[j]])
}
calCeofMat <- function(data, prior){
ngroup <- length(unique(data$groups))  
xbar <- by(data[,c(2,4)], data$groups, colMeans)
n <-by(data[,c(2,4)], data$groups, nrow) 
S <- by(data[,c(2,4)], data$groups, var)
Spool <- matrix(0,2,2)
for(i in 1:ngroup)
  Spool <- Spool+S[[i]]*(n[i]-1)
Spool <- Spool/(sum(n-1))
ceofMat <- matrix(0, ngroup, 3)
for(i in 1:nrow(ceofMat)){
ceofMat[i,] <- c(-0.5*t(xbar[[i]])%*%solve(Spool)%*%xbar[[i]]+log(prior[i]),
                 t(xbar[[i]])%*%solve(Spool)) 
}
ceofMat
}
# coef includes intercept, coefficient for x1, coefficient for x2 (y) 
# get the y value when we know the intercept, coefficients, and x
yval <- function(coef, x)
{
  -coef[1]/coef[3] - coef[2]*x/coef[3] 
}

# the discriminant function
dhat <- function(coef, x1,x2)
{
  cbind(1,x1,x2)%*%coef
}


# draw the boundary between groups g1 and g2
drawline <- function(g1, g2, ngroup, ceofMat, x1){
y1  <- yval(ceofMat[g1,]-ceofMat[g2,], x1)
sel <- rep(TRUE,length(x1))
for(i in 1:ngroup)
{
  if(i!=g1 && i!=g2)
sel <- sel & (dhat(ceofMat[g1,], x1,y1) >= dhat(ceofMat[i,], x1,y1))    
}
lines(x1[sel],y1[sel], lwd=2)
}

# draw all the boundaries between groups 
drawBoundaries <- function(data, prior)
{
x1range <- range(data[,2])
x1points  <- seq(x1range[1], x1range[2], length.out = 200)
ceofMat <- calCeofMat(data, prior)
plot(data[,c(2,4)], col = as.integer(data$groups), pch = as.integer(data$groups))
ngroup <- length(unique(data$groups))  
for(i in 1:(ngroup-1))
  for(j in (i+1):ngroup)
  drawline(i,j, ngroup,ceofMat,x1points)
}
drawBoundaries(t11_5, c(1/3, 1/3, 1/3))
```



e)
```{r}
z <- lda(groups ~ sepal_width + petal_width, data=t11_5, prior=c(1/3, 1/3, 1/3))
conf_matrix <- table((predict(z, newdata=t11_5[,c(2,4)]))$class, t11_5$groups)
APER = (conf_matrix[2,3] + conf_matrix[3,2])/150
APER

z1 <- lda(groups ~ sepal_width + petal_width, data=t11_5, prior=c(1/3, 1/3, 1/3), CV=T)
conf_matrix_h <- table(z1$class, t11_5$groups)
AER <- (conf_matrix_h[2,3] + conf_matrix_h[3,2])/150
AER


```
\[ APER = \frac{1+4}{150} = \frac{1}{30} = 0.033
\]

\[ AER = \frac{2+4}{150} = \frac{6}{150} = 0.040\]


## Question 11.32
a)
```{r}
t11_8 <- read.table("T11-8.DAT.txt", header = F, col.names=c("group", "AFH_activity", "AFH_antigen"))
plot(as.matrix(t11_8[,2:3]), xlab="", ylab="", col = as.integer(t11_8$group), pch= as.integer(t11_8$group))
```
It is fair to assume a bivariate normal distribution because the data points is elliptical shape.

b)
```{r}
xbar1 <- by(t11_8[,2:3], t11_8$group, colMeans)
n1 <-by(t11_8[,2:3], t11_8$group, nrow) 
S1 <- by(t11_8[,2:3], t11_8$group, var)
Spool = (S1[[1]]*(n1[[1]]-1) + S1[[2]]*(n1[[2]]-1))/sum(n1-1)
m_hat <- 0.5*t(xbar1[[1]] - xbar1[[2]])%*%solve(Spool)%*%(xbar1[[1]]+xbar1[[2]])
cat(t(xbar1[[1]]-xbar1[[2]])%*%solve(Spool), -m_hat)

zz <- lda(group ~ AFH_activity + AFH_antigen, data=t11_8, CV=T)

table(zz$class, t11_8$group)
```

We allocate x to population 1 if 
\[19.319x_{1} - 17.12424x_{2} + 3.559472 \ge 0\]
Otherwise, to population 2.

\[ AER = \frac{4+7}{75} = \frac{11}{75} \]

c)
```{r}
new_cases <- matrix(c(-.112,-.279,-0.059,-.068,0.064,0.012,-.043, -.052, -.05, -.098, -.094, -.113, -.123, -.143, -.011, -.037, -.210, -.090, -.126, -.019), 10, 2, byrow=T)

cbind(new_cases, rep(1, 10)) %*% t(cbind(t(xbar1[[1]]-xbar1[[2]])%*%solve(Spool), -m_hat))
```
The list is the 10 linear discrimate functions for the 10 cases. Since all prior probabilities for both populations are equal, $$ ln(\frac{p_i}{p_j}) = 0 $$. All cases have linear discriminate value greater than or equal to 0. We classify the 10 cases under population 1, noncarriers.


d)
```{r}
cbind(new_cases, rep(1, 10)) %*% t(cbind(t(xbar1[[1]]-xbar1[[2]])%*%solve(Spool), -m_hat)) - rep(log(1/3), 10)
```
The above values are calculated from the formula given below.
We classify x to population 1, noncarriers if
\[ (\overline{x}_{1} - \overline{x}_{2})'S^{-1}_{pooled}x - \frac{1}{2}(\bar{x}_{1}-\bar{x}_{2})'S^{-1}_{pooled}(\bar{x}_{1} + \bar{x}_{2}) - \ln(\frac{p_{2}}{p_{1}}) \ge 0   \]
And population 2, otherwise.
We classify the ten cases under population 1, noncarriers, because all the values are greater than 0.



