---
title: "stat445_a3"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## STAT 445 Assignment 3


There are 42 observations from a bivariate normal distribution. The sample mean is given by $\bar{x} = \begin{bmatrix} 0.0564 \\ 0.603 \end{bmatrix}$. The sample variance is given by  $S = \begin{bmatrix} 0.0144 && 0.0117 \\ 0.0117 && 0.0146 \end{bmatrix}$ .
```{r}
S = matrix(c(0.0144, 0.0117, 0.0117, 0.0146), 2, 2)
S
xbar = matrix(c(0.564, 0.603), 2, 1)
xbar
```

## Question 1

a) Find the 95% confidence region for the mean vector $\mu = \begin{bmatrix} \mu_{1} \\ \mu_{2} \end{bmatrix}$.

```{r}
#install.packages("ellipse")
library(ellipse)
```


```{r}
plot(ellipse(S/42, centre=xbar, t=sqrt(((2*41)/40)*qf(0.95, 2, 40) )), type="l", xlim=c(0.51, 0.630), ylim=c(0.5, 0.70))
abline(h=0.565)
abline(h=0.641)
abline(v=0.527)
abline(v=0.601)

abline(v=0.521, lty=2)
abline(v=0.607, lty=2)
abline(h=0.560, lty=2)
abline(h=0.646, lty=2)


```

## Question 2
a)
```{r}
data <- read.table("~/Documents/SFU/STAT445/a1/fat_dat.txt", quote="\"", comment.char="")
fat_dat <- data[,10:19]
n <- nrow(fat_dat)
p <- ncol(fat_dat)
F <- qf(0.95, n, n-p)
cbind(colMeans(fat_dat) - sqrt((p*(n-1)/(n-p))*F)*sqrt(diag(var(fat_dat))/n),colMeans(fat_dat) + sqrt((p*(n-1)/(n-p))*F)*sqrt(diag(var(fat_dat))/n))

```


b)

```{r}
variance <- var(fat_dat[,c(2,3)])
means <- colMeans(fat_dat[,c(2,3)])
plot(ellipse(variance/n, centre=means, t=sqrt(((p*(n-1)/(n-p))*qf(0.95, p, n-p))))
     ,type="l")
```


c) Bonferroni Confidence Intervals for the 

```{r}
circData <- fat_dat
m <- ncol(circData)
t2 <- -qt(0.05/(2*m), df=n-1, lower.tail=TRUE)
intervals <- rbind(colMeans(circData) - t2*sqrt(diag(var(circData))/n), colMeans(circData) + t2*sqrt(diag(var(circData))/n))
intervals

```

d)
```{r}
chestInt <- intervals[,2]
abInt <- intervals[,3]
plot(ellipse(variance/n, centre=means, t=sqrt(((p*(n-1)/(n-p))*qf(0.95, p, n-p)))), type="l")
abline(v=chestInt[1])
abline(v=chestInt[2])
abline(h=abInt[1])
abline(h=abInt[2])
```

e) The confidence interval for mean chest circumference - mean abdomen circumference is (7.44485, 9.091658).
```{r}
m_new = m + 1
b <- matrix(0, 10, 1)
b[2] <- 1
b[3] <- -1
t_new <- -qt(0.05/(2*m_new), df=n-1)
cbind(t(b) %*% colMeans(circData) - t_new * sqrt((t(b) %*% var(circData) %*% b)/n),t(b) %*% colMeans(circData) + t_new * sqrt((t(b) %*% var(circData) %*% b)/n))
```

